Double Standards in Judging Collective Action\thanks{}

### 2023-04-19 Documents/manuscript.tex
switch to Quarto

\documentclass[twocolumn, 11pt, letterpaper]{article}

% Font
\usepackage[MnSymbol]{mathspec}
\setmainfont{Minion Pro}
\setmathsfont(Digits,Latin,Greek)
[Numbers={Lining,Proportional}]{Minion Pro}
\usepackage{microtype}

% Format
\usepackage[letterpaper, margin = 1in]{geometry}
\setlength{\columnsep}{18pt}
\setcounter{secnumdepth}{0}
\usepackage[compact]{titlesec}
\titleformat*{\section}{\centering\normalfont\Large\bfseries}
\titleformat*{\subsection}{\raggedright\normalfont\normalsize\bfseries}
\titleformat*{\subsubsection}{\raggedright\normalfont\normalsize\bfseries\itshape}
\usepackage{authblk}
\renewcommand\Affilfont{\small}

% References/Footnotes
\newcommand{\refsection}{\section{References}}

@@

% Front matter
\title{ Double Standards in Judging Collective Action\thanks{} }
\author[1]{Nils K. Reimer}
\affil[1]{University of California, Santa Barbara, United States}
\author[2]{Marija Brankovi\'{c}}
\affil[2]{Singidunum University and University of Belgrade, Serbia}
\author[3]{Iniobong Essien}
\affil[3]{Leuphana University of L\"{u}neburg, Germany}
\author[4]{Jin X. Goh}
\affil[4]{Colby College, United States}
\author[5]{S\'{e}bastien Goudeau}
\affil[5]{University of Poitiers, France}
\author[6]{N\'{o}ra A. Lantos}
\affil[6]{E\"{o}tv\"{o}s Lor\'{o}nd University, Hungary}
\author[7]{Jenny Veldman}
\affil[7]{NYU Abu Dhabi, United Arab Emirates}

@@

\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
      \noindent Collective action is a powerful force driving social
change but often sparks contention about what actions are acceptable
means to effect social change. In five studies (total \emph{N} = 2,979),
we investigated double standards in judging collective action---that is,
whether observers will judge the same protest actions as more acceptable
if the protesters are ingroup members and their cause aligns with the
ingroup's interests (\emph{identity-based double standards}) or if the
protesters' cause aligns with the observers' ideological positions
(\emph{ideology-based double standards}). In two studies, we developed
an instrument of 25 controversial protest actions, based in item
response theory, to measure where people draw the line between
acceptable and unacceptable forms of collective action. In three
preregistered experiments, we used this instrument to test for double
standards in judging collective action for workers' rights in the United
Kingdom and for or against defunding the police and for or against
restricting abortion in the United States. We found evidence for
ideology-based, but not for identity-based, double standards:
Participants judged the same protest actions to be more acceptable when
the protesters' cause aligned with their own ideological position---but
showed no consistent ingroup bias when judging collective action by
ingroup and outgroup members. Our findings have theoretical and
practical implications for understanding the often divided response to
prominent social movements.
    \end{abstract}
  \end{@twocolumnfalse}
  % \renewcommand{\abstractname}{Statement of Relevance}
  % \begin{@twocolumnfalse}
  %   \maketitle
  %   \begin{abstract}
  %   \noindent 
  %   \end{abstract}
  % \end{@twocolumnfalse}
]

{
  \renewcommand{\thefootnote}%
    {\fnsymbol{footnote}}
  \footnotetext[1]{M.B.--J.V. are listed in alphabetical order. This research was supported by a Seedcorn Research Grant from the European Association of Social Psychology.}
}

\noindent Protest movements often spark contention about what actions
are acceptable means of protest. Reactions to Black American athletes
kneeling during the national anthem to protest racist police violence
are a case in point. Kneeling during the national anthem is not violent,
disruptive, or illegal. And yet, only 29\% of White Americans (compared
to 66\% of Black Americans) and 11\% of Republicans (compared to 59\% of
Democrats) considered it appropriate for Black athletes to protest in
this way (YouGov, 2017). Indeed, Black athletes who had participated in
the protests faced a backlash in public opinion (Lacina, 2020). This and
other examples suggest that \emph{who} the protesters are and
\emph{what} they are protesting influences how acceptable observers
judge their actions to be.

>>>>

\documentclass[12pt, letterpaper]{article}

% Font
\usepackage{amsmath, amssymb}
\usepackage{lmodern}
\usepackage{microtype}

% Format
\usepackage[letterpaper, margin = 1in]{geometry}
\setcounter{secnumdepth}{0}
\setlength{\parindent}{0.5in}
\usepackage[compact]{titlesec}
\titleformat*{\section}{\centering\normalfont\normalsize\bfseries}
\titleformat*{\subsection}{\raggedright\normalfont\normalsize\bfseries}
\titleformat*{\subsubsection}{\raggedright\normalfont\normalsize\bfseries\itshape}
\usepackage{indentfirst}

% References/Footnotes
\newcommand{\refsection}{\newpage \section{References}}
\usepackage[hang, flushmargin]{footmisc}

@@

% Line spacing
\usepackage[doublespacing]{setspace}

@@

\section{Abstract}

\noindent Collective action is a powerful force driving social change
but often sparks contention about what actions are acceptable means to
effect social change. In five studies (total \emph{N} = 2,979), we
investigated double standards in judging collective action---that is,
whether observers will judge the same protest actions as more acceptable
if the protesters are ingroup members and their cause aligns with the
ingroup's interests (\emph{identity-based double standards}) or if the
protesters' cause aligns with the observers' ideological positions
(\emph{ideology-based double standards}). In two studies, we developed
an instrument of 25 controversial protest actions, based in item
response theory, to measure where people draw the line between
acceptable and unacceptable forms of collective action. In three
preregistered experiments, we used this instrument to test for double
standards in judging collective action for workers' rights in the United
Kingdom and for or against defunding the police and for or against
restricting abortion in the United States. We found evidence for
ideology-based, but not for identity-based, double standards:
Participants judged the same protest actions to be more acceptable when
the protesters' cause aligned with their own ideological position---but
showed no consistent ingroup bias when judging collective action by
ingroup and outgroup members. Our findings have theoretical and
practical implications for understanding the often divided response to
prominent social movements.

\textit{Keywords:} collective action, double standards, social identity
theory, system justification theory, Black Lives Matter

\section{ Double Standards in Judging Collective Action }

Protest movements often spark contention about what actions are
acceptable means of protest. Reactions to Black American athletes
kneeling during the national anthem to protest racist police violence
are a case in point. Kneeling during the national anthem is not violent,
disruptive, or illegal. And yet, only 29\% of White Americans (compared
to 66\% of Black Americans) and 11\% of Republicans (compared to 59\% of
Democrats) considered it appropriate for Black athletes to protest in
this way (YouGov, 2017). Indeed, Black athletes who had participated in
the protests faced a backlash in public opinion (Lacina, 2020). This and
other examples suggest that \emph{who} the protesters are and
\emph{what} they are protesting influences how acceptable observers
judge their actions to be.
